---
title: "CH4 analyses"
author: "Abby Lewis"
date: "2026-01-20"
output: html_document
---

## Overview

This file contains code to analyze and visualize the effects of soil warming on methane emissions, including changes in the proportional contribution of extreme fluxes (i.e., top 5%, 1%, and 0.2%). Code chunks should be run in order.

### Table of contents: 

-   Step 1: Load data and packages
-   Step 2: Calculate summary statistics
-   Step 3: Plot data summary (Figure 2 and variants)
-   Step 4: Analyze the contribution of extreme fluxes (Figure 3 and variants)
-   Step 5: Simulate low-frequency sampling to assess bias (Figure 4)
-   Step 6: Daytime and WL sampling biases (SI)
-   Step 7: Analyze residuals for annual and extreme fluxes


## Step 1: Load data and packages

```{r setup}
# Load packages and functions
library(tidyverse)
source("R/hm_style.R")
source("R/plot_all_data.R")
source("R/make_gam_label.R")
source("R/make_lm_label.R")
source("R/calc_bias.R")

# Load data
ch4 <- read_csv(here::here("01_Data", "GENX_flux_data_interpolated_with_drivers.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels),
         date = as_date(DateTime_EST)
  )

soil_temp <- read_csv(here::here("01_Data", "Soil_temp.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels))

smartx <- read_csv(here::here("01_Data", "SMARTX CH4 and CO2 fluxes_2016-2025.csv")) %>%
  mutate(date = as_date(Date, format = "%m/%d/%Y"),
         CH4_umol_m2_h = `CH4 flux (umol/m2/d)`/24)
```


## Step 2: Calculate summary stats

```{r summary_stats}
# Percent gapfilled
ch4 %>%
  mutate(filled_ch4 = ifelse(is.na(CH4_umol_m2_h_not_gapfilled) & !is.na(CH4_umol_m2_h), T, F),
         filled_gpp = ifelse(is.na(GPP_umol_m2_h_not_gapfilled) & 
                               !is.na(GPP_umol_m2_h) &
                               !GPP_umol_m2_h == 0, 
                             T, F)) %>%
  group_by(Chamber) %>%
  summarize(
    pct_ch4 = round(sum(filled_ch4) / sum(!is.na(CH4_umol_m2_h)) * 100, 1),
    pct_gpp = round(sum(filled_gpp) / sum(!is.na(GPP_umol_m2_h)) * 100, 1)
  ) 

# Number of days in analysis
n_days <- as.numeric(difftime(max(ch4$DateTime_EST),
  min(ch4$DateTime_EST),
  units = "days"
))
```


## Step 3: Plot data summary (Figure 2 and variants)

```{r fig_2}
# Standardize x limits
annual_xlim <- range(
  ch4$date[!is.na(ch4$CH4_umol_m2_h)],
  na.rm = TRUE
)

# Set focal dates
start_date <- as_date("2025-03-27")
end_date   <- as_date("2025-04-06")
start_datetime <- as_datetime(paste0(start_date, " 00:00:00"), tz = "EST")
end_datetime <- as_datetime(paste0(end_date, " 23:59:00"), tz = "EST")
focal_days <- seq.Date(start_date, end_date, by = "day")

# Process data
annual_ch4_df <- ch4 %>%
  group_by(date, Chamber) %>%
  summarize(
    CH4_umol_m2_h = mean(CH4_umol_m2_h, na.rm = TRUE),
    Depth_cm = mean(Depth_cm, na.rm = TRUE),
    GPP_umol_m2_h = mean(GPP_umol_m2_h, na.rm = TRUE),
    Reco_umol_m2_h = mean(Reco_umol_m2_h, na.rm = TRUE),
    AirTemp_C = mean(AirTemp_C, na.rm = TRUE),
    Salinity = mean(Salinity, na.rm = TRUE),
    .groups = "drop"
  )

annual_ST_df <- soil_temp %>%
  mutate(date = as_date(DateTime_EST)) %>%
  group_by(date, Chamber) %>%
  summarize(
    Temp_C = mean(SoilTemp_C, na.rm = TRUE),
    .groups = "drop"
  )

focal_ch4_df <- ch4 %>%
  filter(date %in% focal_days)

focal_ST_df <- soil_temp %>%
  mutate(date = as_date(DateTime_EST)) %>%
  filter(date %in% focal_days)

# Plot annual
annual <- make_annual_plot(
  data = annual_ch4_df, 
  x = "date", y = "CH4_umol_m2_h", colour = "Chamber",
  expression(atop(CH[4] ~ "flux","(µmol" ~ m^-2 ~ h^-1 * ")"))
)

annual_ST <- make_annual_plot(
  annual_ST_df,
  "date", "Temp_C", "Chamber",
  expression(atop("Soil temp.", "(ºC)")),
  zero_line = NULL
)

annual_AT <- make_annual_plot(
  annual_ch4_df,
  "date", "AirTemp_C", 
  ylabel = expression(atop("Air temp.", "(ºC)"))
)

annual_wl <- make_annual_plot(
  annual_ch4_df,
  "date", "Depth_cm",
  ylabel = expression(atop("Water level", "(cm)"))
)

annual_sal <- make_annual_plot(
  annual_ch4_df,
  "date", "Salinity",
  ylabel = expression(atop("Salinity", "(PPT)")),
  zero_line = NULL
)

annual_GPP <- make_annual_plot(
  annual_ch4_df,
  "date", "GPP_umol_m2_h", "Chamber",
  ylabel = expression(atop("GPP", "(µmol/m2/h)"))
)

annual_Reco <- make_annual_plot(
  annual_ch4_df,
  "date", "Reco_umol_m2_h", "Chamber",
  ylabel = expression(atop("Reco", "(µmol/m2/h)"))
)

# Plot focal
focal_ch4 <- make_focal_plot(
  focal_ch4_df,
  "DateTime_EST", "CH4_umol_m2_h", "Chamber",
  expression(atop(CH[4] ~ "flux","(µmol" ~ m^-2 ~ h^-1 * ")"))
)

focal_ST <- make_focal_plot(
  focal_ST_df,
  "DateTime_EST", "SoilTemp_C", "Chamber",
  expression(atop("Soil temp.", "(ºC)")),
  zero_line = NULL
)

focal_AT <- make_focal_plot(
  focal_ch4_df,
  "DateTime_EST", "AirTemp_C",
  ylabel = expression(atop("Air temp.", "(ºC)")),
  zero_line = NULL
)

focal_wl <- make_focal_plot(
  focal_ch4_df,
  "DateTime_EST", "Depth_cm",
  ylabel = expression(atop("Water level", "(cm)"))
)

focal_GPP <- make_focal_plot(
  focal_ch4_df,
  "DateTime_EST", "GPP_umol_m2_h", "Chamber",
  ylabel = expression(atop("GPP", "(µmol/m2/h)"))
)

focal_Reco <- make_focal_plot(
  focal_ch4_df,
  "DateTime_EST", "Reco_umol_m2_h", "Chamber",
  ylabel = expression(atop("Reco", "(µmol/m2/h)"))
)

focal_sal <- make_focal_plot(
  focal_ch4_df,
  "DateTime_EST", "Salinity",
  ylabel = expression(atop("Salinity", "(PPT)")),
  zero_line = NULL
)

# Combine annual
annual_panel <- patchwork::wrap_plots(
  annual +
    ggtitle("Seasonal variability (daily mean)") +
    guides(color = guide_legend(ncol = 4)) +
    top_theme +
    panel_theme +
    theme(legend.position = "bottom"),

  annual_ST + top_theme + panel_theme + theme(legend.position = "none"),
  annual_wl + panel_theme + theme(legend.position = "none"),

  ncol = 1,
  design = "
    A
    A
    B
    C
  "
)

# Combine focal
focal_panel <- patchwork::wrap_plots(
  focal_ch4 +
    ggtitle("Short-term variability") +
    guides(color = guide_legend(ncol = 4)) +
    top_theme +
    panel_theme +
    theme(legend.position = "bottom"),

  focal_ST + top_theme + panel_theme + theme(legend.position = "none"),
  focal_wl + panel_theme + theme(legend.position = "none"),

  ncol = 1,
  design = "
    A
    A
    B
    C
  "
)

# Combine both
comb <- patchwork::wrap_plots(
  annual_panel,
  focal_panel,
  ncol = 2,
  widths = c(2, 1),
  guides = "collect"
) +
  patchwork::plot_annotation(
    tag_levels = list(c("a", "","","b")),
    theme = theme(
      legend.position = "bottom",
      legend.justification = "left",
      legend.title.position = "top"
    )
  ) &
  theme(
    plot.tag = element_text(face = "bold")
  )
comb

ggsave(plot = comb, "../03_Figures/Manuscript/Figure 2 - Data drivers.jpg", dpi = 300, width = 7, height = 5)

# Make a smaller version for presentations
patchwork::wrap_plots(
  make_annual_plot(
    data = annual_ch4_df, 
    x = "date", y = "CH4_umol_m2_h", colour = "Chamber",
    expression(atop(CH[4] ~ "flux","(µmol" ~ m^-2 ~ h^-1 * ")")),
    annotate = F
  ) +
    ggtitle("Seasonal variability (daily mean)") +
    guides(color = guide_legend(ncol = 4)) +
    top_theme +
    panel_theme +
    theme(legend.position = "bottom"),
  
  make_annual_plot(
    annual_ST_df,
    "date", "Temp_C", "Chamber",
    expression(atop("Soil temp.", "(ºC)")),
    zero_line = NULL, annotate = F
  ) +
    top_theme +
    panel_theme +
    theme(legend.position = "none"),
  
  ncol = 1, design = c("A
                        A
                        B")
) +
  patchwork::plot_layout(guides = "collect") +
  patchwork::plot_annotation(
    tag_levels = list(c("a", "", "", "b")),
    theme = theme(
      legend.position = "bottom",
      legend.justification = "left",
      legend.title.position = "top"
    )
  ) &
  theme(plot.tag = element_text(face = "bold"))

ggsave("../03_Figures/Presentation/Figure 2 - Data drivers pres.jpg", dpi = 300, width = 7, height = 5)

# Make an even smaller version for presentations
ch4 %>%
  filter(Chamber == "Ch. 12 (+5.7 ºC)") %>%
  ggplot(aes(x = DateTime_EST, y = CH4_umol_m2_h)) +
  geom_line()+
  xlab("Date")+
  guides(color = guide_legend(ncol = 4)) +
  theme_hm +
  theme(legend.position = "none",
        axis.title.y = element_blank()) +
  scale_x_datetime(
    date_breaks = "1 month",
    date_labels = "%b"
  )

ggsave("../03_Figures/Presentation/Figure 2 - Data pres tiny.jpg", dpi = 300, width = 4, height = 3)

# Make an even smaller version for grants
grant <- ch4 %>%
  filter(Chamber == "Ch. 8 (+3.3 ºC)",
         DateTime_EST >= "2025-10-01",
         DateTime_EST < "2025-11-01") %>%
  ggplot(aes(x = DateTime_EST, y = CH4_umol_m2_h)) +
  geom_line()+
  guides(color = guide_legend(ncol = 4)) +
  theme_hm +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  scale_x_datetime(
    breaks = as_datetime(c("2025-10-01 00:00:00", "2025-10-15 00:00:00", "2025-11-01 00:00:00")),
    labels = c("Oct 1", "Oct 15", "Nov 1")
  )

#Ch8 month 9

ggsave(plot = grant, "../03_Figures/Presentation/Figure 2 - Data pres tiny (grant).jpg", dpi = 300, width = 2, height = 1)

# Make an expanded version for the SI
annual_panel <- patchwork::wrap_plots(
  annual +
    ggtitle("Seasonal variability (daily mean)") +
    guides(color = guide_legend(ncol = 4)) +
    top_theme +
    panel_theme +
    theme(legend.position = "bottom"),

  annual_ST + top_theme + panel_theme + theme(legend.position = "none"),
  annual_AT + top_theme + panel_theme + theme(legend.position = "none"),
  annual_GPP + top_theme + panel_theme + theme(legend.position = "none"),
  annual_Reco + top_theme + panel_theme + theme(legend.position = "none"),
  annual_wl + top_theme + panel_theme + theme(legend.position = "none"),
  annual_sal + panel_theme + theme(legend.position = "none"),

  ncol = 1,
  design = "
    A
    A
    B
    C
    D
    E
    F
    G
  "
)

# Combine focal
focal_panel <- patchwork::wrap_plots(
  focal_ch4 +
    ggtitle("Short-term variability") +
    guides(color = guide_legend(ncol = 4)) +
    top_theme +
    panel_theme +
    theme(legend.position = "bottom"),

  focal_ST + top_theme + panel_theme + theme(legend.position = "none"),
  focal_AT + top_theme + panel_theme + theme(legend.position = "none"),
  focal_GPP + top_theme + panel_theme + theme(legend.position = "none"),
  focal_Reco + top_theme + panel_theme + theme(legend.position = "none"),
  focal_wl + top_theme + panel_theme + theme(legend.position = "none"),
  focal_sal + panel_theme + theme(legend.position = "none"),

  ncol = 1,
  design = "
    A
    A
    B
    C
    D
    E
    F
    G
  "
)

# Combine both
comb <- patchwork::wrap_plots(
  annual_panel,
  focal_panel,
  ncol = 2,
  widths = c(2, 1),
  guides = "collect"
) +
  patchwork::plot_annotation(
    tag_levels = list(c("a", "","","","","","","b")),
    theme = theme(
      legend.position = "bottom",
      legend.justification = "left",
      legend.title.position = "top"
    )
  ) &
  theme(
    plot.tag = element_text(face = "bold")
  )
comb

ggsave("../03_Figures/SI/SI - Data drivers expanded.jpg", dpi = 300, width = 7, height = 10)

ch4 %>%
  mutate(DateTime_EST = as.POSIXct(DateTime_EST)) %>%
  ggplot(aes(x = DateTime_EST, y = CH4_umol_m2_h)) +
  scale_x_datetime(
    date_breaks = "1 month",
    date_labels = "%b"
  ) +
  theme_hm +
  theme(
    axis.title.x = element_blank()
  ) +
  labs(y = expression(atop(CH[4] ~ "flux","(µmol" ~ m^-2 ~ h^-1 * ")"))) +
  geom_hline(
    yintercept = 0,
    color = "grey50"
  ) +
  geom_line(
    aes(color = Chamber),
    linewidth = 0.5
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  geom_point(data = smartx %>%
               filter(!substr(Plot, 1,1) == "6",
                      year(date) == 2025,
                      month(date) > 3) %>%
               mutate(DateTime_EST = as.POSIXct(as_datetime(paste0(date, " 00:00:00")))),
             size = 2)

ggsave("../03_Figures/Presentation/Figure 2 - SMARTX comparison.jpg", 
       dpi = 300, width = 5.5, height = 3)
```


## Step 4: Analyze the contribution of extreme fluxes (Figure 3 and variants)

```{r fig_3}
# Only use days that have data from all 12 chambers
hm_df <- ch4 %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(Chamber)) == 12)

# Calculate the contribution of extreme fluxes
hot_moments <- hm_df %>%
  group_by(Chamber) %>%
  # Identify whether fluxes exceed each threshold
  mutate(
    hot_moment_quant_998 = ifelse(CH4_umol_m2_h >=
      quantile(CH4_umol_m2_h,
        probs = 0.998, na.rm = TRUE
      ),
    "hot", "not"
    ),
    hot_moment_quant_99 = ifelse(CH4_umol_m2_h >=
      quantile(CH4_umol_m2_h,
        probs = 0.99, na.rm = TRUE
      ),
    "hot", "not"
    ),
    hot_moment_quant_95 = ifelse(CH4_umol_m2_h >=
      quantile(CH4_umol_m2_h,
        probs = 0.95, na.rm = TRUE
      ),
    "hot", "not"
    )
  ) %>%
  # Calculate sum of fluxes that exceed each threshold
  # and convert to growing season total (mol/m2/y from umol/m2/h)
  summarize(
    n = sum(!is.na(CH4_umol_m2_h)),
    annual_flux = sum(CH4_umol_m2_h, na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_998 = sum(CH4_umol_m2_h[hot_moment_quant_998 == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_99 = sum(CH4_umol_m2_h[hot_moment_quant_99 == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_95 = sum(CH4_umol_m2_h[hot_moment_quant_95 == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    min = min(DateTime_EST),
    max = max(DateTime_EST)
  ) %>%
  mutate( # Add temperature treatment
    Temp_C = sub(" ºC", "", str_extract(as.character(Chamber), "-*[0-9]+\\.*[0-9]* ºC")),
    Temp_C = as.numeric(sub("mb", "0", Temp_C))
  )

write_csv(hot_moments %>%
            select(Chamber,
                   annual_flux,
                   hot_sum_99
                   ), 
          "annual_totals.csv") #Used in a wavelet fig

# Generate panel labels
annual_label <- make_gam_label(
  data = hot_moments,
  xvar = "Temp_C",
  yvar = "annual_flux"
)

# Plot overall growing season totals
annual <- hot_moments %>%
  ggplot(aes(x = Temp_C, y = annual_flux)) +
  geom_smooth(method = "gam",
              formula = y ~ s(x, k = 3),
              method.args = list(method = "REML"),
              color = "black") +
  geom_point(aes(fill = Chamber),
    color = "black", shape = 21,
    stroke = 0.5, size = 2
  ) +
  geom_text(
    aes(x = -Inf, y = Inf, label = annual_label$label),
    hjust = 0, vjust = 1.5,
    inherit.aes = FALSE,
    size = 2.5, data = data.frame()
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Growing season", "emissions (mol" ~ m^-2 ~ yr^-1 * ")"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  theme(
    legend.position = "bottom",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2))

# Calculate summary stats about the contribution of extreme events
top_pct <- hot_moments %>%
  mutate(
    hm_pct_998 = hot_sum_998 / annual_flux * 100,
    hm_pct_99 = hot_sum_99 / annual_flux * 100,
    hm_pct_95 = hot_sum_95 / annual_flux * 100
  ) %>%
  select(Temp_C, Chamber, hm_pct_998, hm_pct_99, hm_pct_95) %>%
  pivot_longer(c(hm_pct_95, hm_pct_99, hm_pct_998)) %>%
  mutate(
    name = factor(
      name,
      levels = c("hm_pct_95", "hm_pct_99", "hm_pct_998"),
      labels = c(
        "Top 5% of measurements",
        "Top 1% of measurements",
        "Top 0.2% of measurements"
      )
    )
  ) %>%
  group_by(name) %>%
  mutate(
    model = list(
      mgcv::gam(
        value ~ s(Temp_C, k = 3),
        data = cur_data(),
        method = "REML"
      )
    ),
    p      = summary(model[[1]])$s.pv,
    r2     = summary(model[[1]])$r.sq,
    median = median(value)
  ) %>%
  ungroup()

# Format summary stats for plot
label_df <- top_pct %>%
  group_by(name) %>%
  summarise(
    p = first(p),
    r2 = first(r2),
    median = first(median),
    label = paste0(
      "  Median = ", round(median, 1), " %",
      "\n  R² = ", round(r2, 2),
      "\n  p = ", round(p, 3)
    )
  )

# Plot the contribution of extreme events to total emissions
top_pct_fig <- top_pct %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(
    method = "gam",
    formula = y ~ s(x, k = 3),
    method.args = list(method = "REML"),
    color = "black",
    data = ~ subset(.x, p < 0.05)
  ) +
  geom_point(aes(fill = Chamber),
    color = "black", shape = 21,
    stroke = 0.5, size = 2
  ) +
  geom_text(
    data = label_df,
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.2,
    inherit.aes = FALSE, size = 2.5
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Percent of growing season", "emissions"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4)) +
  facet_wrap(~name, scales = "free_y")

# Combine growing season sums with extreme events
comb <- ggpubr::ggarrange(annual + theme(legend.position = "right"),
  top_pct_fig + theme(legend.position = "none"),
  nrow = 2, ncol = 1,
  labels = "auto"
)

# Save
ggsave(
  comb,
  filename = "../03_Figures/Manuscript/Figure 3 - Top percent.jpg",
  dpi = 300, width = 6, height = 5
)

# Format 1% panel for presentations
top_pct_pres <- top_pct %>%
  filter(name == "Top 1% of measurements") %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(
    method = "gam",
    formula = y ~ s(x, k = 3),
    method.args = list(method = "REML"),
    color = "black",
    data = ~ subset(.x, p < 0.05)
  ) +
  geom_point(aes(fill = Chamber),
    color = "black", shape = 21,
    stroke = 0.5, size = 2
  ) +
  geom_text(
    data = label_df %>%
      filter(name == "Top 1% of measurements"),
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.5,
    inherit.aes = FALSE, size = 4
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Top 1% of fluxes", "(% of growing season emissions)"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 3))

# Save presentation version
ggsave(
  top_pct_pres,
  filename = "../03_Figures/Presentation/Figure 3 - Top percent pres.jpg",
  dpi = 300, width = 4, height = 5
)

# Format growing season total and 1% panel for presentations
comb_pres <- patchwork::wrap_plots(
  annual + guides(fill = guide_legend(ncol = 3)),
  top_pct_pres,
  nrow = 1, ncol = 2,
  guides = "collect", axes = "collect_x"
) &
  theme(legend.position = "bottom")

ggsave(
  comb_pres,
  filename = "../03_Figures/Presentation/Figure 3 - Top percent pres comb.jpg",
  dpi = 300, width = 8, height = 4
)
```


## Step 5: Simulate low-frequency sampling to assess bias (Figure 4)

```{r fig_4}
# Set simulation specs
reps <- 10000
set.seed(4700)
intervals <- c("daily", "twice weekly", "weekly", "fortnightly", "monthly")

# Add labels for day, month, week, and fortnight
df <- ch4 %>%
  mutate(
    day = yday(date),
    month = month(date),
    week = week(date),
    fortnight = week %/% 2
  )

# Run simulations
df_out <- purrr::map(intervals, calc_bias) %>% # See calc_bias.R
  bind_rows()

# Summarize outputs
to_plot <- df_out %>%
  group_by(Chamber, rep, interval, time) %>%
  summarize(
    mean_hourly = sum(flux * n / sum(n), na.rm = TRUE),
    tot_hourly = mean_hourly / 1e6 * (n_days * 24), # Growing season total (simulated)
    .groups = "drop"
  ) %>%
  left_join(
    hot_moments %>% rename(tot_real = annual_flux), # Add totals (calculated high-frequency)
    by = "Chamber"
  ) %>%
  mutate(percent_bias = tot_hourly / tot_real * 100 - 100)

# Calculate medians
to_plot_median <- to_plot %>%
  group_by(Chamber, interval, time, Temp_C) %>%
  summarize(
    median_bias = median(percent_bias, na.rm = TRUE),
    .groups = "drop"
  )

# Plot medians by time interval
sims_full_day <- to_plot_median %>%
  filter(time == "all") %>%
  mutate(
    interval = factor(interval, levels = intervals),
    int_num = as.numeric(interval)
  ) %>%
  ggplot(aes(color = Chamber, x = int_num, y = median_bias)) +
  geom_hline(yintercept = 0, color = "grey85") +
  geom_line(size = 0.5) +
  geom_point(aes(fill = Chamber), shape = 21, size = 2, stroke = 0.5) +
  scale_x_continuous(
    breaks = 1:length(intervals),
    labels = intervals
  ) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  theme_hm +
  theme(legend.title.position = "top") +
  ylab("Manual sampling bias (%)") +
  xlab("Simulated sampling interval") +
  guides(fill = guide_legend(ncol = 2)) +
  theme(axis.text.x = element_text(
    hjust = 1,
    angle = 45
  ))

# Plot monthly medians by temperature
# First, calculate statistics for plot
sims_label <- make_gam_label(
  data = to_plot_median %>%
    filter(
      interval == "monthly",
      time == "all"
    ),
  xvar = "Temp_C",
  yvar = "median_bias"
)

# Generate plot
sims_month <- to_plot_median %>%
  filter(
    interval == "monthly",
    time == "all"
  ) %>%
  ggplot(aes(x = as.numeric(Temp_C), y = median_bias, group = "group")) +
  geom_hline(yintercept = 0, color = "grey85") +
  geom_smooth(method = "gam",
              formula = y ~ s(x, k = 3),
              method.args = list(method = "REML"),
              color = "black") +
  geom_point(aes(fill = Chamber), shape = 21, size = 2, stroke = 0.5) +
  geom_text(
    data = sims_label,
    aes(x = -Inf, y = -Inf, label = label),
    hjust = 0, vjust = -0.5,
    inherit.aes = FALSE, size = 2.5
  ) +
  scale_x_continuous(
    limits = c(-0.5, 6.5),
    expand = c(0, 0)
  ) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  theme_hm +
  theme(legend.title.position = "top") +
  ylab("Manual sampling bias (%)") +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 2))

# Plot full distribution
sims_dist <- to_plot %>%
  filter(
    interval == "monthly",
    time == "all"
  ) %>%
  ggplot(aes(x = as.numeric(Temp_C), y = percent_bias)) +
  geom_hline(yintercept = 0, color = "grey85") +
  # Without quantiles for shading
  geom_violin(aes(color = Chamber, fill = Chamber),
    position = "identity", alpha = 0.1, width = 0.5
  ) +
  # With 25 and 75%
  geom_violin(aes(color = Chamber), fill = NA,
    quantiles = c(0.25, 0.75), quantile.linetype = "11",
    position = "identity", width = 0.5
  ) +
  # With median
  geom_violin(aes(color = Chamber), fill = NA,
    quantiles = c(0.5), quantile.linetype = "solid",
    position = "identity", width = 0.5
  ) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  theme_hm +
  scale_x_continuous(
    limits = c(-0.5, 6.5),
    expand = c(0, 0)
  ) +
  theme(legend.title.position = "top") +
  ylab("Manual sampling bias (%)") +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 2))

# Combine plots
with_dist2 <- patchwork::wrap_plots(
  a = sims_month +
    guides(fill = guide_legend(ncol = 1)) +
    ylab("Monthly sampling bias\n(median %)") +
    xlab("Temperature treatment (ºC above ambient)"),
  b = sims_dist + theme(legend.position = "none") +
    ylab("Monthly sampling bias (%)"),
  c = sims_full_day +
    ylab("Sampling bias\n(median %)") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(hjust = 0.5, angle = 0)
    ),
  d = patchwork::guide_area(),
  guides = "collect",
  axes = "collect_x",
  design = c("AAABBB
              AAABBB
              AAABBB
              CCCCDD
              CCCCDD")
) +
  patchwork::plot_annotation(
    tag_levels = "a"
  ) &
  theme(plot.tag = element_text(face = "bold"))

# Visualize and save
with_dist2
ggsave(with_dist2,
  filename = "../03_Figures/Manuscript/Figure 4 - Simulations all.jpg",
  dpi = 300, width = 6, height = 5
)
```


## Step 6: Daytime and WL sampling biases (SI)

```{r fig_4}
# Set simulation specs
reps <- 10000
set.seed(4700)
intervals <- c("monthly")

# Add labels for day, month, week, and fortnight
df <- ch4 %>%
  mutate(
    day = yday(date),
    month = month(date),
    week = week(date),
    fortnight = week %/% 2,
    DateTime_local = with_tz(DateTime_EST, tzone = "America/New_York"))

# Run simulations
df_out <- calc_bias("monthly", only_time = F) %>% # See calc_bias.R
  bind_rows()

# Summarize outputs
to_plot <- df_out %>%
  group_by(Chamber, rep, interval, time) %>%
  summarize(
    mean_hourly = sum(flux * n / sum(n), na.rm = TRUE),
    tot_hourly = mean_hourly / 1e6 * (n_days * 24), # Growing season total (simulated)
    .groups = "drop"
  ) %>%
  left_join(
    hot_moments %>% rename(tot_real = annual_flux), # Add totals (calculated high-frequency)
    by = "Chamber"
  ) %>%
  mutate(percent_bias = tot_hourly / tot_real * 100 - 100)

# Calculate medians
to_plot_median <- to_plot %>%
  group_by(Chamber, interval, time, Temp_C) %>%
  summarize(
    median_bias = median(percent_bias, na.rm = TRUE),
    .groups = "drop"
  )

# Stats
to_plot_median %>%
  group_by(interval, time) %>%
  summarize(median = median(median_bias))

# Plot medians by time interval
sims_multiple_biases <- to_plot_median %>%
  mutate(
    time = factor(time, 
                  levels = c("all", "daytime", "waterlevel"),
                  labels = c("Fully random\nsampling times", 
                             "Only daytime\nsampling", 
                             "Only sampling\nwhen water level\nis at or below the\nmarsh surface")),
  ) %>%
  ggplot(aes(x = time, y = median_bias)) +
  geom_hline(yintercept = 0, color = "grey85") +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(fill = Chamber), shape = 21, size = 2, stroke = 0.5, width = 0.2) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  coord_cartesian(xlim = c(0.5, 3.5), expand = c(left = F, right = F))+
  theme_hm +
  theme(legend.title.position = "top") +
  ylab("Bias of monthly sampling (%)") +
  xlab("Simulated sampling bias") +
  guides(fill = guide_legend(ncol = 2)) +
  theme(axis.title.x = element_blank(),
        legend.box.margin = margin(t = -25))

sims_multiple_biases

ggsave(sims_multiple_biases,
  filename = "../03_Figures/SI/SI - Multiple biases.jpg",
  dpi = 300, width = 3.5, height = 4
)
```


## Step 7: Analyze residuals for annual and extreme fluxes

```{r resids}
# Load driver data
rtk <- read_csv(here::here("01_Data", "Chamber_elevation.csv")) %>%
  select(-Latitude_DD, -Longitude_DD) %>%
  mutate(Type = "") %>%
  pivot_longer(-c(Chamber, Type))
agb <- read_csv(here::here("01_Data", "Aboveground_biomass.csv"),
  show_col_types = F
) %>%
  mutate(Type = "Aboveground~biomass~(g/m^2)") %>%
  pivot_longer(-c(Chamber, Type))
stems <- read_csv(here::here("01_Data", "SCAM_Stem_measurements.csv")) %>%
  mutate(Type = "italic('S. americanus')~stem~traits") %>%
  pivot_longer(-c(Chamber, Type)) %>%
  group_by(Chamber, Type, name) %>%
  summarize(value = median(value))
counts <- read_csv(here::here("01_Data", "SCAM_Stem_counts.csv")) %>%
  mutate(Type = "italic('S. americanus')~stem~traits") %>%
  pivot_longer(-c(Chamber, Type))

# Combine and format drivers
drivers <- rtk %>%
  bind_rows(agb) %>%
  bind_rows(stems) %>%
  bind_rows(counts) %>%
  mutate(name = case_match(
    name,
    "Elev_m" ~ "Elevation~(m)",
    "SCAM_g_m2" ~ "C[3]",
    "TotalC4_g_m2" ~ "C[4]",
    "TotalAboveground_g_m2" ~ "Total",
    "SCAM_Height_cm" ~ "Height~(cm)",
    "SCAM_Width_mm" ~ "Width~(mm)",
    "SCAM_Stem_count" ~ "Count"
  ))

# Calculate residuals
resids <- hot_moments %>%
  mutate(
    hm_pct_998 = hot_sum_998 / annual_flux * 100,
    hm_pct_99 = hot_sum_99 / annual_flux * 100,
    hm_pct_95 = hot_sum_95 / annual_flux * 100
  ) %>%
  select(Chamber, Temp_C, hm_pct_998, hm_pct_99, hm_pct_95, annual_flux) %>%
  pivot_longer(-c(Chamber, Temp_C), names_to = "HM_variable") %>%
  group_by(HM_variable) %>%
  mutate(resid = mgcv::gam(value ~ s(Temp_C, k = 3))$residuals) %>%
  mutate(HM_variable = replace_values(
    HM_variable,
    "hm_pct_99" ~ "'Top 1%'",
    "hm_pct_998" ~ "'Top 0.2%'",
    "hm_pct_95" ~ "'Top 5%'",
    "annual_flux" ~ "Total"
  )) %>%
  select(-value)

for_plot <- resids %>%
  left_join(drivers, relationship = "many-to-many") %>%
  mutate(Chamber = factor(Chamber, levels = chamber_labels))

# Generate labels
labs <- make_lm_label(for_plot, "value", "resid",
  group_var = "HM_variable", group_var2 = "name"
) %>%
  left_join(for_plot %>%
    select(name, Type) %>%
    distinct())

# Plot
pollock <- for_plot %>%
  ggplot(aes(x = value, y = resid)) +
  geom_point(aes(color = Chamber)) +
  theme_hm +
  scale_color_manual(values = color.gradient, name = "Chamber") +
  xlab("value") +
  ylab("Residual") +
  geom_text(
    aes(x = x_pos, y = y_pos, label = label_simple),
    hjust = 0, vjust = 1.1,
    inherit.aes = FALSE,
    size = 3, data = labs, parse = T
  ) +
  ggh4x::facet_nested("Methane~flux~metric" + HM_variable ~ Type + name,
    scales = "free", switch = "x",
    labeller = label_parsed,
    nest_line = element_line(linewidth = 0.2)
  ) +
  theme(
    strip.placement = "outside",
    axis.title.x = element_blank()
  )

# Save
ggsave(pollock,
  file = here::here("03_Figures", "SI", "SI - HM_residuals.jpg"),
  width = 10, height = 5
)
```