---
title: "Wavelet analysis"
author: "Abby Lewis"
date: "2024-08-22"
output: html_document
---

## Overview

This file contains code to run wavelet analyses and assess the effect of warming on each scale of wavelet variability. Code chunks should be run in order.

### Table of contents:

-   Step 1: Load data and packages
-   Step 2: Run wavelet analysis
-   Step 3: Plot variance by temperature (Figure 5)
-   Step 4: Visualize reconstructed wavelet outputs (SI)
-   Step 5: Assess Spearman correlation with drivers (Figure 6)
-   Step 6: Analyze residuals for each scale of variability (SI)


## Step 1: Load data and packages

```{r setup}
# Load packages
library(tidyverse)
library(data.table)
library(waveslim)
library(DescTools)
source("R/wavelet_functions.R")
source("R/make_gam_label.R")
source("R/make_lm_label.R")
source("R/hm_style.R")

# Load data
ch4 <- read_csv(here::here("processed_data", "L2- partitioned_and_gap_filled.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST")) %>%
  mutate(Chamber = paste0(location, Fluxing_Chamber))

ch4_sums <- ch4 %>%
  mutate(Chamber = paste0(location, Fluxing_Chamber))%>%
  group_by(Chamber) %>%
  summarize(median = median(CH4_umol_m2_h, na.rm = T))

chamber_labels = unique(ch4$Chamber)
```


## Step 2: Run wavelet analysis

```{r run_wavelets}
# SPECIFICATIONS
# Choose drivers
drivers <- c(
  "AirTemp_C",
  "GPP_umol_m2_h",
  "Reco_umol_m2_h", 
  "SlrFD_W_Avg"
)

# Set how drivers should be referenced in plots
driver_labels <- c(
  AirTemp_C = "Air temperature",
  GPP_umol_m2_h = "GPP",
  GPP_unfilled = "GPP",
  Reco_umol_m2_h = "Reco",
  SlrFD_W_Avg = "Solar radiation"
)

# Set driver colors
driver_colors <- c(
  AirTemp_C = "#ff8b85",
  GPP_umol_m2_h = "#3BAA5D",
  GPP_unfilled = "#3BAA5D",
  Reco_umol_m2_h = "#264653",
  SlrFD_W_Avg = "gold"
)

# Run separate wavelet analyses across all treatments and sites
dfs <- map(
  c("CH4_umol_m2_h", "CO2_umol_m2_h", drivers),
  ~ double_map(
    treatment = unique(ch4$Chamber),
    ch4 = ch4,
    var_name = .x,
    timestep_s = 6750
  )
) %>%
  bind_rows()

# Re-introduce NAs
nas_to_add <- ch4 %>%
  mutate(
    na_ch4 = ifelse(is.na(CH4_umol_m2_h_not_gapfilled), T, F),
    na_co2 = ifelse(is.na(CO2_umol_m2_h_not_gapfilled), T, F),
    na_gpp = ifelse(is.na(GPP_umol_m2_h_not_gapfilled), T, F)
  ) %>%
  select(DateTime_EST, Chamber, na_ch4, na_gpp, na_co2) %>%
  distinct()

# Group wavelet scales into interpretable bins
grouped_dfs <- dfs %>%
  mutate(
    name = as.character(name),
    num_value = as.numeric(gsub("[^0-9]", "", name)),
    grouped_scale = case_when(
      grepl("days", name) & num_value < 30 ~ "Weekly",
      grepl("hours", name) & num_value < 10 ~ "Hourly",
      grepl("hours", name) & num_value >= 10 ~ "Daily",
      TRUE ~ "Seasonal"
    )
  ) %>%
  select(-num_value) %>%
  group_by(grouped_scale, Chamber, DateTime_EST, var_name) %>%
  summarize(
    value = sum(value, na.rm = T),
    reconstructed = sum(reconstructed, na.rm = T),
    .groups = "drop"
  )

# Make a wider version
discrete_scales <- grouped_dfs %>%
  pivot_wider(names_from = var_name, values_from = c(value, reconstructed)) %>%
  rename_with(~ gsub("value_", "", .x)) %>%
  left_join(nas_to_add) %>%
  mutate(
    CH4_unfilled = ifelse(na_ch4, NA, CH4_umol_m2_h),
    CO2_unfilled = ifelse(na_co2, NA, CO2_umol_m2_h),
    GPP_unfilled = ifelse(na_gpp, NA, GPP_umol_m2_h),
    Chamber = factor(Chamber,
      levels = chamber_labels
    ),
    grouped_scale = factor(grouped_scale,
      levels = c("Hourly", "Daily", "Weekly", "Seasonal")
    )
  ) %>%
  select(-na_ch4, -na_gpp, -na_co2)

# Separate explanatory variables
discrete_scales_exp <- discrete_scales %>%
  dplyr::select(-matches("CO2_unfilled|CO2_umol_m2_h|CH4_unfilled|CH4_umol_m2_h|GPP_unfilled|recon")) %>%
  pivot_longer(
    cols = -c(grouped_scale, Chamber, DateTime_EST),
    names_to = "exp_var", values_to = "exp_var_value"
  ) %>%
  filter(!is.na(exp_var_value)) %>%
  rename(grouped_scale_exp = grouped_scale)
```


## Step 3: Plot variance by temperature (Figure 5)

```{r plot_variance}
# Calculate the total variance at each scale (CH4)
grouped_pct_var <- grouped_dfs %>%
  filter(var_name %in% c("CH4_umol_m2_h", "CO2_umol_m2_h")) %>%
  group_by(grouped_scale, Chamber, var_name) %>%
  summarize(
    var = var(value, na.rm = T),
    .groups = "drop"
  ) %>%
  group_by(Chamber, var_name) %>%
  mutate(tot_var = sum(var, na.rm = T)) %>%
  ungroup() %>%
  mutate(
    Chamber = factor(Chamber,
      levels = chamber_labels
    ),
    grouped_scale = factor(grouped_scale,
      levels = c("Hourly", "Daily", "Weekly", "Seasonal")
    )
  )

grouped_pct_var %>%
  mutate(location = sub("[0-9]+", "", Chamber),
         ch_n = sub("[a-z]+", "", Chamber),
         grouped_scale = factor(grouped_scale,
      levels = levels(grouped_scale)
    )
         ) %>%
  ggplot(aes(x = grouped_scale, y = var/tot_var, group = ch_n, color = ch_n))+
  geom_line()+
  ylab("Proportion of total variance")+
  facet_grid(var_name~location)+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm
```


## Step 5: Assess Spearman correlation with drivers (Figure 6)

```{r cors}
# Calculate correlations
spearman_corrs <- discrete_scales %>%
  dplyr::select(c(grouped_scale, Chamber, DateTime_EST,
                  CH4_umol_m2_h, CO2_umol_m2_h)) %>%
  pivot_longer(c(CH4_umol_m2_h, CO2_umol_m2_h), 
               names_to = "response_var", 
               values_to = "response_var_value") %>%
  left_join(discrete_scales_exp, relationship = "many-to-many") %>%
  filter(
    !is.na(exp_var_value),
    !is.na(response_var_value)
  ) %>%
  group_by(Chamber, grouped_scale, exp_var, grouped_scale_exp, response_var) %>%
  summarize(
    spearman_rho = cor(exp_var_value, response_var_value,
      method = "spearman",
      use = "complete.obs"
    ),
    spearman_rho2 = spearman_rho^2,
    abs_rho = abs(spearman_rho),
    .groups = "drop"
  ) %>%
  filter(
    grouped_scale == grouped_scale_exp
  )

# Plot correlations
spear <- spearman_corrs %>%
  mutate(location = sub("[0-9]+", "", Chamber)) %>%
  ggplot(aes(
    x = grouped_scale,
    y = spearman_rho,
    color = exp_var,
    group = exp_var
  )) +
  geom_hline(color = "grey50", yintercept = 0) +
  geom_line(alpha = 0.1, aes(group = paste0(Chamber, exp_var))) +
  geom_line(
    data = . %>% group_by(exp_var, grouped_scale, response_var) %>%
      summarize(spearman_rho = mean(spearman_rho)),
    linewidth = 1.5
  ) +
  scale_x_discrete(expand = c(0, 0.4)) +
  xlab("Wavelet scale of variation") +
  ylab("Spearman correlation (ρ)") +
  scale_color_manual(
    values = driver_colors[drivers],
    labels = driver_labels[drivers],
    name = "Explanatory variable"
  ) +
  guides(color = guide_legend(ncol = 2)) +
  theme_hm +
  facet_grid(response_var~location)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Visualize and save
spear
ggsave(spear, filename = "Figures/Spearman correlations.png", width = 4, height = 3.5)

# Now try not grouping
ungrouped_dfs <- dfs %>%
  mutate(
    name = as.character(name),
    grouped_scale = name,
    num_value = as.numeric(gsub("[^0-9]", "", name)),
    text = gsub("[0-9]+ ", "", name),
    text = factor(text, levels = c("hours", "days"))
  ) %>%
  group_by(grouped_scale, Chamber, DateTime_EST, var_name, text, num_value) %>%
  summarize(
    value = sum(value, na.rm = T),
    reconstructed = sum(reconstructed, na.rm = T),
    .groups = "drop"
  )

levels <- ungrouped_dfs %>%
  select(grouped_scale, text, num_value) %>%
  distinct() %>%
  arrange(text, num_value)

#trimmed_levels <- levels %>%
#  filter(num_value < 50)

# Make a wider version
discrete_scales_ungrouped <- ungrouped_dfs %>%
  pivot_wider(names_from = var_name, values_from = c(value, reconstructed)) %>%
  rename_with(~ gsub("value_", "", .x)) %>%
  left_join(nas_to_add) %>%
  mutate(
    CH4_unfilled = ifelse(na_ch4, NA, CH4_umol_m2_h),
    CO2_unfilled = ifelse(na_co2, NA, CO2_umol_m2_h),
    GPP_unfilled = ifelse(na_gpp, NA, GPP_umol_m2_h),
    Chamber = factor(Chamber,
      levels = chamber_labels
    ),
    grouped_scale = factor(grouped_scale,
      levels = unique(levels$grouped_scale)
    )
  ) %>%
  filter(!is.na(grouped_scale)) %>%
  select(-na_ch4, -na_co2, -na_gpp, -text, -num_value)

# Separate explanatory variables
discrete_scales_exp_ungrouped <- discrete_scales_ungrouped %>%
  dplyr::select(-matches("CO2_unfilled|CH4_unfilled|CH4_umol_m2_h|GPP_unfilled|recon")) %>%
  pivot_longer(
    cols = -c(grouped_scale, Chamber, DateTime_EST),
    names_to = "exp_var", values_to = "exp_var_value"
  ) %>%
  filter(!is.na(exp_var_value)) %>%
  rename(grouped_scale_exp = grouped_scale)

# Calculate correlations
spearman_corrs_ungrouped <- discrete_scales_ungrouped %>%
  dplyr::select(c(grouped_scale, Chamber, DateTime_EST, CH4_umol_m2_h)) %>%
  left_join(discrete_scales_exp_ungrouped, relationship = "many-to-many") %>%
  filter(
    !is.na(exp_var_value),
    !is.na(CH4_umol_m2_h)
  ) %>%
  group_by(Chamber, grouped_scale, exp_var, grouped_scale_exp) %>%
  summarize(
    spearman_rho = cor(exp_var_value, CH4_umol_m2_h,
      method = "spearman",
      use = "complete.obs"
    ),
    spearman_rho2 = spearman_rho^2,
    abs_rho = abs(spearman_rho),
    .groups = "drop"
  ) %>%
  filter(
    grouped_scale == grouped_scale_exp
  )

# Plot correlations
spear_ungrouped <- spearman_corrs_ungrouped %>%
  ggplot(aes(
    x = grouped_scale,
    y = spearman_rho,
    color = exp_var,
    group = exp_var
  )) +
  geom_hline(color = "grey50", yintercept = 0) +
  geom_line(alpha = 0.1, aes(group = paste0(Chamber, exp_var))) +
  geom_line(
    data = . %>% group_by(exp_var, grouped_scale) %>%
      summarize(spearman_rho = mean(spearman_rho)),
    linewidth = 1.5
  ) +
  scale_x_discrete(expand = c(0, 0.4)) +
  xlab("Wavelet scale of variation") +
  ylab("Spearman correlation (ρ)") +
  scale_color_manual(
    values = driver_colors[drivers],
    labels = driver_labels[drivers],
    name = "Explanatory variable"
  ) +
  guides(color = guide_legend(ncol = 2)) +
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Visualize and save
spear_ungrouped
ggsave(spear_ungrouped, filename = "Figures/Spearman correlations ungrouped.png", width = 4, height = 4)
```

Summary without grouping

```{r}
ungrouped_pct_var <- ungrouped_dfs %>%
  filter(var_name %in% c("CH4_umol_m2_h", "CO2_umol_m2_h")) %>%
  group_by(grouped_scale, Chamber, var_name) %>%
  summarize(
    var = var(value, na.rm = T),
    .groups = "drop"
  ) %>%
  group_by(Chamber, var_name) %>%
  mutate(tot_var = sum(var, na.rm = T)) %>%
  ungroup() %>%
  mutate(
    Chamber = factor(Chamber,
      levels = chamber_labels
    )
  )

levels <- ungrouped_dfs %>%
  select(grouped_scale, text, num_value) %>%
  distinct() %>%
  arrange(text, num_value)

ungrouped_pct_var %>%
  mutate(location = sub("[0-9]+", "", Chamber),
         ch_n = sub("[a-z]+", "", Chamber),
    grouped_scale = factor(grouped_scale,
      levels = unique(levels$grouped_scale)
    )
         ) %>%
  ggplot(aes(x = grouped_scale, y = var/tot_var, group = ch_n, color = ch_n))+
  geom_line()+
  ylab("Proportion of total variance")+
  facet_grid(var_name~location)+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

ungrouped_pct_var %>%
  mutate(location = sub("[0-9]+", "", Chamber),
         ch_n = sub("[a-z]+", "", Chamber),
    grouped_scale = factor(grouped_scale,
      levels = unique(levels$grouped_scale)
    )
         ) %>%
  ggplot(aes(x = grouped_scale, y = var, group = ch_n, color = ch_n))+
  geom_line()+
  ylab("Total variance")+
  facet_grid(var_name~location, scales = "free_y")+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

